---
title: "MSB106 - Assignment"
author: "Sindre H. Øveraas, Alen Colakovic, Mona Lisa Jones & Sebastian M. Fløysand"
format:
 pdf:
  fontsize: "10"
  documentclass: article
  margin-left: 2cm
  margin-right: 2cm
papersize: A4
editor: visual
---

```{r, message=FALSE}
library(rgdal)
library(dplyr)
library(RSQLite)
library(sf)
library(tidyverse)
library(readr)
library(ggplot2)
library(kableExtra)
```

```{r, message=FALSE}
NOR_CBD <- read_csv("NOR_CBD.csv")
Dist_CBD_Dentist <- read_csv("Dist_CBD_Dentist.csv")
Zonal_CBD <- read_csv("Zonal_CBD.csv")
Dist_Mal_Dentist <- read.csv("Dist_Mal_Dentist.csv")
NOR_OSM_SHOP_MAL <- read.csv("NOR_OSM_SHOP_MAL.csv")
Zonal_Den <- read.csv("Zonal_Den.csv")
NOR_KOMM <- read_csv("NOR_KOMM.csv")
NOR_KOMM_2 <- read_csv("NOR_KOMM_2.csv")
```

```{r}
Commune_Data <- inner_join(NOR_KOMM, NOR_KOMM_2, by = "kommunenummer") |>
  select(kommunenummer, HubName, HubDist, Turnover_capita_retail_Omsetning) |>
  rename("knr" = kommunenummer, "DistMal" = HubDist, "HubNameMal" = HubName, "Turnover_capita_retail" = Turnover_capita_retail_Omsetning)
```

```{r}
Dentist_Data <- inner_join(Dist_Mal_Dentist, Dist_CBD_Dentist, by = "fid") |>
  select(fid, Juridisk.n, Antall.ans, Sum.Drifts, Sum.salgsi, Driftsresu.y, osm_id.y, latlong.y, HubName.y, HubDist.y, HubName.x, HubDist.x) |>
  rename("DistMal" = HubDist.x, "DistCBD" = HubDist.y, "HubNameMal" = HubName.x, "HubNameCBD" = HubName.y, "Driftsres" = Driftsresu.y, "osm_id" = osm_id.y, "latlong" = latlong.y)
```

## Introduction

Our main task in this assignment is to perform an analysis of geospatial determinants of firm activity.
More specifically we are to focus on the Norwegian dental industry in this regard, and see how geospatial determinants such as distances to shopping malls and CBDs (Central Business Districts), as well as population density can determine dental businesses income and general financial operations.
As an example, central questions in this assignment will be; "Is it more beneficial to be highly centralized in urban areas with high population density and many competitors, or is it a greater advantage to be less centralized to the advantage that the nearest competing company is considerably further away?", "Which determinants appear to be most significant for economic benefit?"

## Theoretical Foundation of Hypothesis

## Data description

```{r}
kable(summary(Commune_Data))
```

```{r}
kable(summary(Dentist_Data[3:6]))
```

```{r}
kable(summary(Dentist_Data[8:12]))
```

```{r}
Commune_Data |>
  ggplot(aes(x = DistMal, y = Turnover_capita_retail, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "Turnover capita in retail per inhabitant") +
  theme_bw()
```

Might be more appropriate to explore the distance from CBDs to Malls compared to turnover capita in retail per inhabitant, rather than the distance from the geographical centroid of Norwegian Communes, because of the tendency of bigger population density around CBDs.

```{r}
Dentist_Data |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r}
Dentist_Data_LIMIT <- Dentist_Data |>
  filter(Sum.salgsi < 100000000, DistCBD < 250,
         Driftsres < 20000000, Driftsres > -10000000, DistMal < 80)
```

```{r}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistCBD, y = Sum.salgsi, colour = DistCBD)) +
  geom_point(lwd = .9) + labs(x = "Distance from CBD", y = "Income") +
  theme_bw()
```

```{r}
Dentist_Data |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

```{r}
Dentist_Data_LIMIT |>
  ggplot(aes(x = DistMal, y = Driftsres, colour = DistMal)) +
  geom_point(lwd = .9) +
  labs(x = "Distance from Mall", y = "EBIT") +
  theme_bw()
```

## Econometric approach

After a quixk view of some of our findings where we used the dentist income (y) and in comparison to the Distance from Mall (x) we already see some signs of a pattern where the income is grater when the dentist is placed closer to the shopping mall and decreases as the distance gets bigger.
However, when the distance reach around 10km from the mall the income starts to increase before it then starts to decrease again around 20km.
The theory from Keynesian suggest that this pattern might take place as the transportation cost for individuals further from the Shopping mall does not increase to much by travelling the extra distance.
Hotelling also describes this case where the consumer are willing to travel further to reach a certain central place.

\

We also find that most dentist are either located close to a City, Central Business District or a shopping mall.

\

Dentist are also located close to echeother making for a sort of cluster, indicating that it might be som advantages beeing located in a cluster.
Taken for granted that or data on Dentist includes both "normal Dentis" and Spescialist, there is reason to believe that it could be an advantage beeing located close together.
As found in Marshalls Theory their ar several advantages by locating in clusters.
Where in this case a "normal dentist" would for example recommend one of their customers to go see a specialist to get braces.
The customer would then probably prefer to go to the closes specialist.
The access to work force in these areas would also be good and wll be whitin the same geographic area.

\
as mentioned earlier, dentis should place them self in the pink area of the Alons model.
however, a dentist beeing placed further out in to other retail agglomeration areas can also be a strategy, but the dentist would then rely on the area having less competition so the rate between competition and population density are proposinal.
Our findings in QGIS shows that in areas where it is a high population density, the number of dentist also is higher.
The user/pasient then has more to choose between and the dentist also has a bigger scale of potential customers.
Where in smaller areas it is less competition and the dentis would appear to have somewhat of a monopoly since the transportation cost to the second closes dentis are noticeable higher.

\

it is important to be aware of the fact that going to the dentist is a public services and most people get assigned to their closest dentis based on their living address.
This also takes away som of the competition.
Another thing that is important to notice is that when passed 18 years of age, you could risk loosing your local public dentist due to lack of capacity.
a suggestion would be that for a dentist to be placed close CBD areas, shopping malls and industry can have a positive effect on revenue since this is areas where most of the population spends their time during their day, either going to work, school and other errand.
So by beeing placed close CBD areas, shopping malls and industry the transportation cost for the customer could be reduced as they are already located in these areas even if they dont live here

\

![](https://lh3.googleusercontent.com/zSwLHynV_h_vnX66jchxqoZgBDHH_LVE5PAOiiSLtM3fgSrVBOPT4lT3G7jP-OlsOwg65bJhrAj9rPTicCosZdGAYHySGihdbgtSzj7UICa8fQfX7oMPtO-xdHGtzP_SnGVXlWLrUUPKDKTWfJN2l00)

\
\

The model shows that moving further away from a CBD and a Mall has a negative effect on te income for dentist.
moving 1 km away from CBD reduce the income by aprox.
528 kr and moving 1 km further away from a shopping mall reduces the income by 25 725 kr.
This tells a story where it is more profitable beeing located close or in a Mal or a CBD where the Mall is the best option out from this model.
The result also indicates that the best placing for a dentis would be in the second circle from the Alonso model as stated earlier, as this is the area where we are most likely to find Shopping malls.

\

The reasoning for a more profitable business close to a CBD can be that this is more crowded areas, gravitation pull.
More people live here and it also contains more business and activities which can be combined with a dentist appointment.
Somewhat of the same pattern goes for shopping malls where customers can combine their dentis visit with shopping afterwards, maybe in a way to threat them self.

\
\
